# Reproducible Research: Peer Assessment 1
```{r setoptions, echo = TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r load_data}
unzip("activity.zip")
activity <- read.csv("activity.csv", sep = ",", header = TRUE, 
                    stringsAsFactors=FALSE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day  
2. Calculate and report the mean and median total number of steps taken per day  
```{r total_daily_steps}
dailysteps <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
hist(dailysteps, 
     col = "red",
     main = "Histogram of the Total Number of Steps Taken Each Day", 
     xlab = "Number of Steps",
     ylab = "Number of Days")

mean(dailysteps)
median(dailysteps)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)  
2. Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?  
```{r daily_pattern}
avgsteps <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
intervals <- activity[activity$date == "2012-10-12", ]$interval
plot(intervals, avgsteps, type = "l",
     main = "Average Number of Steps Taken in 5-minute Intervals",
     xlab = "Interval",
     ylab = "Number of Steps")

intervals[avgsteps == max(avgsteps)]
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset  

2. Devise a strategy for filling in all of the missing values in the dataset -- 
in this case using mean for that 5-minute interval.   

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.  

4. Make a histogram of the total number of steps taken each day and 
Calculate and report the mean and median total number of steps taken per day. 
```{r impute}
dim(activity[!complete.cases(activity), ])[1]

activity$impute <- activity$steps
activity$impute[is.na(activity$impute)] <- avgsteps[match(activity$interval, intervals)]   

dailysteps2 <- tapply(activity$impute, activity$date, sum)
hist(dailysteps2, 
     col = "red",
     main = "Histogram of the Total Number of Steps After Imputing", 
     xlab = "Number of Steps",
     ylab = "Number of Days")

mean(dailysteps2)
median(dailysteps2)
summary(activity)
```
Do these values differ from the estimates from the first part of the assignment?  
- YES  

What is the impact of imputing missing data on the estimates of the total daily number of steps?  
- The density looks more like a normal density with the mean and the median being equal. 

## Are there differences in activity patterns between weekdays and weekends?
Use weekdays() function  

1. Create a new factor variable in the dataset with two levels 
– "weekday" and "weekend" indicating whether a given date 
is a weekday or weekend day.

2. Make a panel plot containing a time series plot (i.e. type = "l") 
of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis).  
```{r weekdays}
library(lattice)
activity$day <- as.factor(ifelse(weekdays(activity$date) %in% 
                                 c("Saturday","Sunday"), "weekend", "weekday"))

activity2 <- split(activity, activity$day)
avg_weekday <- tapply(activity2$weekday$steps, activity2$weekday$interval, 
                      mean, na.rm = TRUE)
avg_weekend <- tapply(activity2$weekend$steps, activity2$weekend$interval, 
                      mean, na.rm = TRUE)

act_weekday <- data.frame(cbind(avg_weekday, intervals, "weekday"), 
                          stringsAsFactors = FALSE)
act_weekend <- data.frame(cbind(avg_weekend, intervals, "weekend"), 
                          stringsAsFactors = FALSE)
names(act_weekday) <- c("steps", "interval", "day")
names(act_weekend) <- c("steps", "interval", "day")

a <- rbind(act_weekday, act_weekend)

a$steps <- as.numeric(a$steps)
a$interval <- as.numeric(a$interval)

par(mfrow = c(2, 1))
plot(intervals, avg_weekday, type = "l",
     main = "Weekday",
     xlab = "Interval",
     ylab = "Number of steps")

plot(intervals, avg_weekend, type = "l",
     main = "Weekend",
     xlab = "Interval",
     ylab = "Number of steps")

xyplot(steps ~ interval | day, data = a, layout = c(1, 2), 
       type = "l",
       xlab = "Interval",
       ylab = "Number of steps")

```
