# Reproducible Research: Peer Assessment 1
```{r setoptions, echo = TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r load_data}
unzip("activity.zip")
activity <- read.csv("activity.csv", sep = ",", header = TRUE, 
                    stringsAsFactors=FALSE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day  
2. Calculate and report the mean and median total number of steps taken per day  
```{r total_daily_steps}
dailysteps <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
hist(dailysteps, 
     col = "red",
     main = "Histogram of the Total Number of Steps Taken Each Day", 
     xlab = "Number of Steps",
     ylab = "Number of Days")

mean(dailysteps)
median(dailysteps)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)  
2. Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?  
```{r daily_pattern}
avgsteps <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
intervals <- activity[activity$date == "2012-10-12", ]$interval
plot(intervals, avgsteps, type = "l",
     main = "Average Number of Steps Taken in 5-minute Intervals",
     xlab = "Interval",
     ylab = "Number of Steps")

intervals[avgsteps == max(avgsteps)]
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset  

2. Devise a strategy for filling in all of the missing values in the dataset -- 
in this case using mean for that 5-minute interval.   

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.  

4. Make a histogram of the total number of steps taken each day and 
Calculate and report the mean and median total number of steps taken per day. 
```{r impute}
dim(activity[!complete.cases(activity), ])[1]

activity$impute <- activity$steps
activity$impute[is.na(activity$impute)] <- avgsteps[match(activity$interval, intervals)]   

dailysteps2 <- tapply(activity$impute, activity$date, sum)
hist(dailysteps2, 
     col = "red",
     main = "Histogram of the Total Number of Steps After Imputing", 
     xlab = "Number of Steps",
     ylab = "Number of Days")

mean(dailysteps2)
median(dailysteps2)
summary(activity)
```
Do these values differ from the estimates from the first part of the assignment?  
- YES  

What is the impact of imputing missing data on the estimates of the total daily number of steps?  
- The density looks more like a normal density with the mean and the median being equal. 

## Are there differences in activity patterns between weekdays and weekends?
